---
title: 'Time Series: Analysis and its applications'
output:
  pdf_document: default
  html_document:
    df_print: paged
---



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

# 1. INTRODUÇÃO

```{r}
library(astsa)
```

# 2. Usando algumas bases de dados do livro 

```{r}
plot(jj, type = "o", ylab="Quarterly Earnings per Share")
```
```{r}
plot(speech)
```
```{r}
WN <- arima.sim(model = list(order = c(0, 0, 0)), n = 200)
```


```{r}
plot.ts(WN,col=4, main="White Noise Series")
```

```{r}
AR<-x <- arima.sim(model = list(ar = c(0.1,0.1)), n = 1000)
plot.ts(WN,col=10, main="ar2 PROCESS")
```
```{r}
acf2(AR)
```

```{r}
ts.sim <- arima.sim(list(order = c(1,1,0), ar = 0.7), n = 200)
ts.plot(ts.sim)
```
```{r}
x<-arima.sim(model = list(order=c(2,0,0), ar=c(1,-0.9)),n=500)
ts.plot(x,col='blue')
```

```{r}
set.seed(154)
```

```{r}
w = rnorm(200,0,1) 
x = cumsum(w)
wd = w +.2 
xd = cumsum(wd)
plot.ts(xd, ylim=c(-5,55), main="random walk")
lines(x)
lines(.2*(1:200), lty="dashed")

```

```{r}
cs <- 2*cos(2*pi*1:500/50 + .6*pi)
w <- rnorm(500,0,1)
par(mfrow=c(3,1), mar=c(3,2,2,1), cex.main=1.5)
plot.ts(cs, main=expression(2*cos(2*pi*t/50+.6*pi)))
plot.ts(cs+w, main=expression(2*cos(2*pi*t/50+.6*pi) + N(0,1)))
plot.ts(cs+5*w, main=expression(2*cos(2*pi*t/50+.6*pi) + N(0,25)))
```
```{r}
acf(speech,250)
```
```{r}
par(mfrow <- c(3,1))
acf(soi, 48, main="Southern Oscillation Index")
acf(rec, 48, main="Recruitment")
ccf(soi, rec, 48, main="SOI vs Recruitment", ylab="CCF")
```

```{r}
ACF  = ARMAacf(ar=c(1.5,-.75), ma=0, 24)[-1]
PACF = ARMAacf(ar=c(1.5,-.75), ma=0, 24, pacf=TRUE)
par(mfrow=c(1,2))
plot(ACF, type="h", xlab="lag", ylim=c(-.8,1)); abline(h=0)
plot(PACF, type="h", xlab="lag", ylim=c(-.8,1)); abline(h=0)
               
```

```{r}
acf2(rec, 48) # will produce values and a graphic
(regr = ar.ols(rec, order=2, demean=FALSE, intercept=TRUE))
regr$asy.se.coef # standard errors of the estimates
```
```{r}
library(astsa)
library(dplyr)
library(glmnet)
set.seed(1)

dados = read.table("http://statweb.stanford.edu/~tibs/ElemStatLearn/datasets/prostate.data")



```


```{r}
tr <- dados$train
```

```{r}
X_tr <- dados[tr, 1:8] %>% as.matrix()
y_tr <- dados$lpsa[tr]
```

```{r}
X_val <- dados[-tr, 1:8] %>% as.matrix()
y_val <- dados$lpsa[-tr]

```

```{r}
ajuste_mq <- glmnet(X_tr, y_tr, alpha = 0, lambda = 0)
predito_mq <- predict(ajuste_mq, newx = X_val)
```



```{r}
cv_ridge <- cv.glmnet(X_tr, y_tr, alpha = 0)
ajuste_ridge <- glmnet(X_tr, y_tr, alpha = 0)
predito_ridge <- predict(ajuste_ridge,s = cv_ridge$lambda.1se,newx = X_val)
```

```{r}
cv_lasso <- cv.glmnet(X_tr, y_tr, alpha = 1)
ajuste_lasso <- glmnet(X_tr, y_tr, alpha = 1)
predito_lasso <- predict(ajuste_lasso,s = cv_lasso$lambda.1se, newx = X_val)

```

```{r}
plot(ajuste_lasso)
plot(cv_lasso)
```
```{r}
print(ajuste_lasso)
```
```{r}
plot(cv_lasso)
```
```{r}
#Gerando e plotando duas séries 
library(ggplot2)
set.seed(12345)

x<-seq(1,10)
y1<- runif(10, min = 7, max =10)
y2<- runif(10, min = 7, max =10)
y3<- runif(10, min = 7, max =10)


df <- data.frame(x,y1,y2,y3)

ggplot(data = df, aes(x = x, y = y1, colour = "1")) +
  geom_line(colour = "red") +
  geom_line(aes(x = x, y = y2, colour = "2"), colour = "green") +
  geom_line(aes(x = x, y = y3, colour = "3"), colour = "blue") +
  labs(title = "TITULO", x = NULL, y = "Temperatura (C)") +
  ylim(0,10)

  





```

```{r}
# Livro de Pedro André Morettin
```


```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\poluicao.csv"

df <-read.csv2(file=pathFile, sep=";",dec=",", header =TRUE)
df
```


```{r}
```
```{r}
library(tidyverse)

sapply(df,typeof)

str(df)
```

```{r}
df$DATA_TS <- as.Date(df$DATA_TS,format="%Y-%m-%d")

str(df)
```


```{r}
df
```
```{r}
library(ggplot2)

p<-ggplot(df, aes(x = DATA_TS, y = PM10)) +
       geom_line()

p<-p+ggtitle("Índice de Poluição Cidade de São Paulo: Concentração Média diária")

p<- p+xlab("'Jan a Dez 97") +ylab("Concentração Média PM10 ")
p
```

```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\temperatura.csv"
df <-read.csv2(file=pathFile, sep=";",dec=",")
df$DATA<- as.Date(paste(df$DATA, "-01", sep=""),'%Y-%m-%d')
str(df)
df
```
```{r}
#df$DATA <- as.Date(df$DATA,format="%Y-%m-%d")

str(df)
```
```{r}
p<-ggplot(df, aes(x = DATA, y = Cananeia)) +
       geom_line()
p<-p+ggtitle("Temperatura Mensal Cananeia")

p<- p+xlab("Jan 1976 a Jan 1985 ") +ylab("Temperatura Mensal Média")
p
```
```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\D-IBV.csv"
df <-read.csv2(file=pathFile, sep=";",dec=".")
df$DATA<- as.Date(df$DATA,'%Y-%m-%d')
str(df)
df
```
```{r}
```


```{r}
p<-ggplot(df, aes(x = DATA, y = df$IBV_D)) +
       geom_line()
p<-p+ggtitle("Indice diário Bovespa")

p<- p+xlab("Agosto de 1998 a Maio de 2001 ") +ylab("IBV.Díario")
p
```
```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\fortaleza.csv"
df <-read.csv2(file=pathFile, sep=";",dec=",")
df$Ano<- as.Date(paste(df$Ano, "-01-01", sep=""),'%Y-%m-%d')
str(df)
df

```
```{r}
p<-ggplot(df, aes(x = Ano, y = df$Fort)) +
       geom_line()
p<-p+ggtitle("Indice diário Bovespa")

p<- p+xlab("Precipitação Pluviométrica Anual- Fortaleza ") +ylab("Precipitação Anual")
p
```
```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\MANCHAS.csv"
df <-read.csv2(file=pathFile, sep=";",dec=",")
df$Ano<- as.Date(paste(df$Ano, "-01-01", sep=""),'%Y-%m-%d')
str(df)
df
```
```{r}
p<-ggplot(df, aes(x = Ano, y = df$manchas)) +
       geom_line()
p<-p+ggtitle("Número médio Anual de Manchas Solares")

p<- p+xlab("Ano") +ylab("Número de Manchas Solares")
p
```
```{r}
pathFile <- "E:\\data\\projeto_latex\\timeseries\\R_CODE\\pibanual.csv"
df <-read.csv2(file=pathFile, sep=";",dec=",")
df$ano<- as.Date(paste(df$ano, "-01-01", sep=""),'%Y-%m-%d')
str(df)
df
```
```{r}
p<-ggplot(df, aes(x = ano, y = df$pib1949)) +
       geom_line()
p<-p+ggtitle("Pib Anual R$- Preços de 1949")

p<- p+xlab("Ano") +ylab("PIB Anual Preços de 1949")
p
```

